A More Complete Algebra for File Synchronization

### 2015-11-12 2015/paper.tex
Description of reconciliation; proof without "fully typed" sequences

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setvx}[1]{\mathcal{V}_{#1}}
\newcommand{\setf}{\setvx{F}}
\newcommand{\setd}{\setvx{D}}
\newcommand{\setb}{\setvx{\empt}}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\toppath}{\mathtt{None}}
\newcommand{\fsbroken}{\mathtt{Broken}}
\newcommand{\FS}{\Phi} % {\mathrm{FS}}
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\pp}{p^\uparrow} % parent path
\newcommand{\np}{p_{\centernot\leftrightarrow}} % path not related to p

@@

\newcommand{\extset}[1]{\wp{#1}}
\newcommand{\orderset}[1]{\vec{\wp}{#1}}

@@

\begin{mydef}{$\orderset{U}$}
We use $\orderset{U}$ to denote the set of sequences that contain all commands in $U$
and represent all orders generated by this algorithm.
At times by $\orderset{U}$ we mean an arbitrary sequence in the set.
\end{mydef}

@@

We have seen that commands can break a filesystem
either because their input type does not match the filesystem at the given path,
or because the filesystem loses its tree property.

\begin{mydef}{Fully typed sequences}
We call a sequence of commands fully typed if whenever it breaks a filesystem
it is explained by a mismatch on the input types of the commands it contains.
% TODO
TODO: define fully typed sequences in terms of the algebra?
\end{mydef}

To aid the construction of fully typed sequences, we reintroduce
two assertion commands that either break the filesystem
or leave it unchanged: $\cbb(p)$ and $\cdd(p)$.
(As $|\setb|=|\setd|=1$, we do not note their exact output value.)
They can be used to test whether $\FS(p)=\empt$ or $\FS(p)=D$.

@@

\begin{itemize}
\item Laws for construction typing:
   \begin{gather*}
            [\cbf(p, F)] \equiv [\cdd(\pp); \cbf(p, F)]\\
            [\cbd(p, D)] \equiv [\cdd(\pp); \cbd(p, D)]
   \end{gather*}
\item Laws for destruction typing:
   \begin{gather*}
            [\cdb(\pp)] \equiv [\cbb(p); \cdb(\pp)]\\
            [\cdf(\pp, F)] \equiv [\cbb(p); \cdf(\pp, F)]
   \end{gather*}
\end{itemize}

@@

\begin{myax}$\cxy(p,V_Y)\ncoworks \czw(p,V_W)$ if $X\neq Z$\end{myax}
% \item Disjoint laws for construction:
%    \begin{gather*}
%    % bd>bf
%             [\cbd(\pp, D); \cbf(p, F)] \ncoworks [\cbf(p, F)]\\
%    % bd>bd
%             [\cbd(\pp, D_1); \cbd(p, D_2)] \ncoworks [\cbd(p, D_2)]\\
%    % fd>bf
%             [\cfd(\pp, D); \cbf(p, F)] \ncoworks [\cbf(p, F)]\\
%    % fd>bd
%             [\cfd(\pp, D_1); \cbd(p, D_2)] \ncoworks [\cbd(p, D_2)]
%    \end{gather*}
% \item Disjoint laws for destruction:
%    \begin{gather*}
%    % fb<db ??
%             [\cfb(p); \cdb(\pp)] \ncoworks [\cdb(\pp)]\\
%    % fb<df ??
%             [\cfb(p); \cdf(\pp, F)] \ncoworks [\cdf(\pp, F)]\\
%    % db<db ??
%             [\cdb(p); \cdb(\pp)] \ncoworks [\cdb(\pp)]\\
%    % db<df ??
%             [\cdb(p); \cdf(\pp, F)] \ncoworks [\cdf(\pp, F)]
%    \end{gather*}

\begin{mylem}
If $A$ and $B$ are minimal sequences, $A\coworks B$,
and there are commands on path $p$ in both $A$ ($\cxy(p, V_Y)$) and $B$ ($\czw(p, V_W)$)
then their input types must match ($X=Z$).
TODO: Can this be derived from the algebra?
\end{mylem}

\begin{mylem}\label{lemma:wrextprefix}
$[S;A]\wrext A$
if $A$ is fully typed and $[S;A]$ is minimal.
TODO: Is this an axiom?
\end{mylem}

\begin{mylem}\label{lemma:ncowprefix}
$A\ncoworks B \Rightarrow [S;A]\ncoworks B$ 
if $A$ is fully typed and $[S;A]$ is minimal.
\end{mylem}
\begin{proof}
This follows from lemma \ref{lemma:wrextprefix} and axiom \ref{axiom:ncowwrnrw}.
\end{proof}

@@

\begin{mylem}
Let $A$ and $B$ be two minimal sequences so that $A\coworks B$.
Then their intersection ($\orderset(A\cap B)$) can also be applied to any filesystem
either can be applied to without breaking it.
\end{mylem}

@@

We show $A$ can be transformed via equivalences so that this would be true
by induction on the number of non-marked commands preceding the last marked command.
If it is 0, no transformation is needed.
For the induction step, select the last marked command preceded by a non-marked command.
If the two commute, reverse them, thereby reducing the number on relevant non-marked commands.
If they do not commute, they must be a construction or destruction pair as we know
that $A$ is minimal and $A\nequiv\cbrk$.

Let the marked command be $\cxy(p, V_Y)$ and the preceding unmarked command in $A$ be $\czw(q, V_W)$.
We have two cases and we will proceed by indirection.

If $B$ has a command on $q$, let it be $\cqr(q, V_R)$. As $A\coworks B$, we know
that $Q=Z$. From lemmas \ref{lemma:neighbour}, \ref{lemma:minextset} and \ref{lemma:minextorder}
we also know that there is a $B^*\equiv B$ where $\cxy$ and $\cqr$ are next to each other,
and from the construction and destruction laws we know that $\cxy$ determines that like
$\czw(q, V_W)$ in $A$, any command on $q$ must precede it in $B$. As $B$ is minimal, $B^*$
contains the same commands, and so it is the $\cqr$ command that precedes $\cxy$ in $B^*$.
Going back to the construction and destruction laws we see that the output type of the first command
is always the same, and it is either $D$ or $\empt$. Therefore $R=W$ and $\cqr(q, V_R)=\czw(q, V_W)$
as $|\setd|=|\setb|=1$. This is a contradiction as $\czw$ was not marked, but we see it must also be in $B$.

If $B$ has no command on $q$, $A$ and $B$ cannot work on the same filesystem,
as the second commands in all construction and destruction laws require a different type of input
as the first commands. This is a contradiction as we supposed $A\coworks B$.

In terms of the algebra, this latter argument could be rephrased as follows.
Split $A$ into $[A_0;\czw(q, V_W);\cxy(p, V_Y);A_1]$.
Add command $\cqq(q)$ before $\cxy(p, V_Y)$ to $B$ according to the construction or destruction typing laws.
This yields $B^*\equiv B$, which we split into $[B_0;\cqq(q);\cxy(p, V_Y);B_1]$.
We know $[\czw(q, V_W);\cxy(p, V_Y)]$ is fully typed (TODO: why?)
and $A$ is minimal, so $[A_0;\czw(q, V_W);\cxy(p, V_Y)]$ is minimal.
We also know $[\cqq(q);\cxy(p, V_Y)]$ is fully typed and $B^*$ is minimal, so $[B_0;\cqq(q);\cxy(p, V_Y)]$ is minimal.
But we know $[\czw(q, V_W);\cxy(p, V_Y)]\ncoworks [\cqq(q);\cxy(p, V_Y)]$
as comparing the construction and destruction laws to their typing counterparts we know that $Z\neq Q$,
and so from lemma \ref{lemma:ncowprefix} and lemma \ref{lemma:ncowpostfix}
we know that $A\ncoworks B^*\equiv B$ and so $A\ncoworks B$ which is a contradiction.
(TODO: Is there a point in proving this this way?)

@@

We now know that $A$ can be rearranged in a way that commands in $A\cap B$
are at the beginning, and therefore $A\wrext \orderset(A\cap B)$ and by symmetry $B\wrext \orderset(A\cap B)$.
\end{proof}

>>>>

% TODO Use cleveref to add prefix to references

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setvx}[1]{\mathcal{V}_{#1}}
\newcommand{\setf}{\setvx{F}}
\newcommand{\setd}{\setvx{D}}
\newcommand{\setb}{\setvx{\empt}}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\toppath}{\mathtt{None}} % return value of parent()
\newcommand{\fsbroken}{\mathtt{Broken}} % broken FS value
\newcommand{\FS}{\Phi} % a filesystem
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\pp}{p^\uparrow} % parent path
\newcommand{\np}{p_{\centernot\leftrightarrow}} % path not related to p

@@

\newcommand{\extset}[1]{\wp({#1})}
\newcommand{\orderset}[1]{\vec{\wp}({#1})}
\newcommand{\ordered}[1]{\vec{\mathbf{o}}({#1})}

@@

\begin{mydef}{$\orderset{U}$}
We use $\orderset{U}$ to denote the set of sequences that contain all commands in $U$
and represent all orders generated by this algorithm.
By $\ordered{U}$ we mean an arbitrary sequence in the set.
\end{mydef}

@@

We reintroduce two assertion commands that either break the filesystem
or leave it unchanged: $\cbb(p)$ and $\cdd(p)$.
(As $|\setb|=|\setd|=1$, we do not note their exact output value.)
They can be used to test whether $\FS(p)=\empt$ or $\FS(p)=D$.

@@

\begin{itemize}
\item Laws for construction assertion:
   \begin{gather*}
            [\cbf(p, F)] \equiv [\cdd(\pp); \cbf(p, F)]\\
            [\cbd(p, D)] \equiv [\cdd(\pp); \cbd(p, D)]
   \end{gather*}
\item Laws for destruction assertion:
   \begin{gather*}
            [\cdb(\pp)] \equiv [\cbb(p); \cdb(\pp)]\\
            [\cdf(\pp, F)] \equiv [\cbb(p); \cdf(\pp, F)]
   \end{gather*}
\end{itemize}

@@

% \begin{myax}$\cxy(p,V_Y)\ncoworks \czw(p,V_W)$ if $X\neq Z$\end{myax}

\begin{mylem}\label{lemma:cowinputmatch}
If $A$ and $B$ are minimal sequences, $A\coworks B$,
and there are commands on path $p$ in both $A$ ($\cxy(p, V_Y)\in A$) and $B$ ($\czw(p, V_W)\in B$)
then their input types must match ($X=Z$).
TODO: Can this be derived from the algebra?
TODO: maybe this is the only thing needed
\end{mylem}

@@

If we start with two copies of the filesystem $\FS$,
and two different sequences are applied to the copies to yield $\FS_A:=A\FS$
and $\FS_B:=B\FS$, then our aim is to define sequences of commands $A^R$ and $B^R$
so that $B^R\FS_A$ and $A^R\FS_B$ would be as close to each other as possible.

We work based on the assumption that to achieve this, we need
to apply to $\FS_B$ the commands that have been applied to $\FS_A$, and \emph{vice versa}.
As some commands may have been applied to both filesystems, our first approximation
is $A^R = \ordered{A\setminus B}$ and $B^R = \ordered{B\setminus A}$.
This, however, will break both filesystems if there have been incompatible updates
in $A$ and $B$. 
Our aim is therefore to provide an algorithm that selects the commands 
$A^R = \ordered{A^{R*}}$  where $A^{R*} \subset A\setminus B$
and $B^R = \ordered{B^{R*}}$ where $B^{R*} \subset B\setminus A$ 
so that $A^R\FS_B\neq\fsbroken$ and $B^R\FS_A\neq\fsbroken$,
and show that these are the longest sequences with this property, that is,
adding any command left out from $A^R$ or $B^R$ will break the filesystems.

We assume that $A$ and $B$ are returned by an update detector,
and so they are minimal and do not contain assertion commands.
We also assume that $\FS_A$ and $\FS_B$ are not broken and therefore $A\coworks B$.

Our reconciliation algorithm is as follows.

\begin{mydef}{Reconciliation}\label{def:reconciliation}
\[ A^R = \ordered{\{c| c\in A\setminus B \wedge \forall c'\in B\setminus A: [c;c']\equiv[c';c]\}} \]

In other words, the sequence $A^R$ we apply to $\FS_B$ are all commands from $A\setminus B$
that commute with every command in $B\setminus A$.
This is equivalent to saying that those commands are selected from $A\setminus B$
whose paths are unrelated to the paths of the commands in $B\setminus A$.
\end{mydef}

Now we prove that $A^R\FS_B\neq\fsbroken$.

\begin{mylem}
Let $A$ and $B$ be two minimal sequences without assertion commands so that $A\coworks B$.
Then their intersection ($\ordered{A\cap B}$) can also be applied to any filesystem
either can be applied to without breaking it.
\end{mylem}

@@

We show $A$ can be transformed via equivalences so that this would be true
by induction on the number of non-marked commands preceding the last marked command.
If it is 0, no transformation is needed.
For the induction step, select the last marked command preceded by a non-marked command.
If the two commute, reverse them, thereby reducing the number on relevant non-marked commands.
If they do not commute, they must be a construction or destruction pair as we know
that $A$ is minimal and $A\nequiv\cbrk$.
Therefore either $q=\parent(p)$ or $p=\parent(q)$.

Let the marked command be $\cxy(p, V_Y)$ and the preceding unmarked command in $A$ be $\czw(q, V_W)$.

If $B$ has a command on $q$, then
from lemma \ref{lemma:neighbour}
we know that there is a $B'$ in $\orderset{B}$ where it is next to $\cxy$,
and from lemmas \ref{lemma:minextset} and \ref{lemma:minextorder} we know that $B'\equiv B$.
From the construction and destruction laws we also know that a command on $q$
must precede $\cxy(p, V_Y)$ just as it does so in $A$.

If $B$ has no command on $q$, then let $B'$ be $B$, but with an extra command added to it just before $\cxy$
according to the construction or destruction assertion laws, from which we know
that $B'\equiv B$.
We also know that $B'$ is still minimal.

We therefore have a minimal $B'$ which has a command on $q$ just before $\cxy(p, V_Y)$ and $B'\equiv B$.
Let this command on $q$ be $\cqr(q, V_R)$.
\begin{gather*}
A = [\cdots; \czw(q, V_W); \cxy(p, V_Y); \cdots] \\
B' = [\cdots; \cqr(q, V_R); \cxy(p, V_Y); \cdots]
\end{gather*}

As $A\coworks B$, 
from lemma \ref{lemma:cowinputmatch}
we know that $Q=Z$.

Going back to the construction and destruction laws we see that the output type of the first command
is always the same, and it is either $D$ or $\empt$. Therefore $R=W$ and $\cqr(q, V_R)=\czw(q, V_W)$
as $|\setd|=|\setb|=1$. 
If $B$ originally had a command on $q$,
this is a contradiction as $\czw$ was not marked, but we see it must also be in $B'$ and therefore in $B$.
If $B$ had no command on $q$,
this is a contradiction because the command added to $B$ is an assertion command, so $Z=Q=R=W$, 
but $A$ contained no assertion commands.

@@

We now know that there is an $A'\equiv A$ in which commands in $A\cap B$
are at the beginning, and therefore 
from axiom \ref{axiom:wrextpostfix}
$A\wrext \ordered{A\cap B}$ and by symmetry $B\wrext \ordered{A\cap B}$.
\end{proof}
